---
title: "Midterm: Physiological Arousal and Political Beliefs"
output:
  pdf_document:
    md_extension: +raw_tex
bibliography: citations.bib
---

```{r setup, include=FALSE}
# Replication of "Physiological Arousal and Political Beliefs"
# Published in Political Psychology (2014)
# Original authors: Jonathan Renshon, Julia Lee, and Dustin Tingley

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stargazer)
library(stringr)
library(bookdown)
library(mediation)
library(memisc)
library(extrafont)
library(rstanarm)
library(janitor)
library(bayesplot)
```

\begin{flushright}
Gabe Walker

March 16, 2019
\end{flushright}

```{r data, echo=FALSE, message=FALSE, error=FALSE}

# Read in the replication file provided on the Harvard Dataverse:
# https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/24318

df <- read_csv("replication_files/anxiety.csv") %>%

Really good to document handling of missing data. Nice.

  # Remove the two individuals who have a lot of missing data.
  # This results in 138 observations as written in the paper.

  filter(!{
    is.na(income) & is.na(education)
  }) %>%

  # Rename and restructure variables for ease of understanding.
  # "video" refers to which video the subject was shown.
  
  Also a great choice. Really helpful for future you and anyone who tries to replicate

  mutate(video = factor(case_when(
    anxcond3 == 0 ~ "relax",
    anxcond3 == 1 ~ "neutral",
    anxcond3 == 2 ~ "anxiety"
  ))) %>%
  dplyr::select(-anxcond3) %>%

  # "anxiety" is a dummy for whether the subject was shown the Cliffhanger clip.

  mutate(anxiety = factor(anxcond)) %>%
  dplyr::select(-anxcond) %>%

  # "skin" is the standardized skin-conductance level within each participant.
  # See the original paper (p. 7) for a detailed description of standardization.
  
  There were 2 measures of skin conductivity, SCDBradVidManipAll_mean and SCDBradSelfReport1_mean.
  This is why your replication of figure two below is slightly off.

  mutate(skin = SCDBradSelfReport1_mean) %>%
  dplyr::select(-SCDBradSelfReport1_mean) %>%

  # "story" is a random dummy for which immigration article the participants read.

  mutate(story = factor(storycond)) %>%
  dplyr::select(-storycond)
```
Nice reformatting/manipulation of the data

### Abstract

I partially replicate "Physiological Arousal and Political Beliefs" by @renshon. The main regression results (Table 1) appear exactly as in the original paper. I have improved the aesthetics of Figures 2 and 3 with a more minimal and unified style. Note that the results of Figure 2 are different than in the original. Though I find significantly different skin-conductance reactivities for each video  treatment group, the overall conclusions are no different. Figure 3 results are the same as in the paper. For the original replication data, see @data.

In my extension I attempt to confirm the main results of the paper using Bayesian linear models. I find that the new models produce findings equivalent to the original ones, which lends credence to the authors' argument that physiological reactivity mediates the relationship between induced anxiety and attitudes toward immigrants.

Great abstract. Thorough and informative.

\vspace{2em}

**Table 1.** Main Results: Physiological reactivity mediates the relationship between induced anxiety and attitudes toward immigrants.

\begin{center}

```{r table1, echo=FALSE, message=FALSE, error=FALSE, results='asis'}
### Table 1.

# Subset the data to only include individuals in the
# "neutral" or "anxiety" video treatment groups.

no_relax <- df %>%
  filter(video != "relax")

# Regress SC reactivity on Anxiety, excluding the relax video group.

mod1 <- lm(skin ~ anxiety, no_relax)

# Regress immigration preferences on anxiety, story, and SC reactivity.

mod2 <- lm(immigration ~ anxiety + story + skin, no_relax)

# Display the results of the two models with
# total number of participants and R^2.

stargazer(mod1, mod2,
  style = "apsr",
  omit.stat = c("adj.rsq", "ll", "F", "ser"),
  covariate.labels = c("Anxiety Manipulation", "Story Condition", "SC Reactivity while answering questions"),
  dep.var.labels = c("SC Reactivity", "Immigration Preferences"),
  column.sep.width = "5pt",
  float = FALSE,
  header = FALSE
)
```
Excellent replication of Table 1. Your footnotes make it clearer than what is in the published paper. Your comments that 
explain what is going on are really helpful.

\end{center}

*Note.* Model (1) shows the effect of the treatment (anxiety) on physiological reactivity while Model (2) shows the effects
of physiological reactivity on immigration preferences, controlling for the story condition. Both models include only
*Neutral* & *Anxiety* conditions (the *Relax* condition is excluded). 

Standard errors in parentheses.

Note that a higher coefficients correspond to *less* favorable views on immigration.

```{r figure2, echo=FALSE, message=FALSE, error=FALSE, cache=TRUE}
### Figure 2.

# Make three data subsets for each treatment group.
# Remove the NA values so that certain functions (e.g., mean()) work.

relax <- df %>%
  filter(video == "relax") %>%
  na.omit()

neutral <- df %>%
  filter(video == "neutral") %>%
  na.omit()

anxiety <- df %>%
  filter(video == "anxiety") %>%
  na.omit()

# Make a plottable dataframe with the means,
# confidence intervals, and counts for the "skin"
# variable for each group

# Find the means of each group, which will
# be used as the center points of the confidence
# intervals.

means <- c(mean(relax$skin), mean(neutral$skin), mean(anxiety$skin))

# Find the confidence interval sizes for each group, i.e.,
# the distance above and below the mean that the confidence
# interval stretches.

ci_relax <- 1.96 * sd(relax$skin) / sqrt(length(relax$skin))
ci_neutral <- 1.96 * sd(neutral$skin) / sqrt(length(neutral$skin))
ci_anxiety <- 1.96 * sd(anxiety$skin) / sqrt(length(anxiety$skin))

# Find the upper and lower bounds for each group.

lower <- c(means[1] - ci_relax, means[2] - ci_neutral, means[3] - ci_anxiety)
upper <- c(means[1] + ci_relax, means[2] + ci_neutral, means[3] + ci_anxiety)

# Put the results in a graphable dataframe.

f2 <- tibble(
  treatment = c("Relax", "Neutral", "Anxiety"),
  mean = means,
  lower = lower,
  upper = upper,
  count = c(nrow(relax), nrow(neutral), nrow(anxiety)),
  x = c(1, 2, 3)
)

# Plot the result. This figure includes a title that
# summarizes the main findings of the figure, color-codes
# the treatment effect, and shows which of the confidence
# intervals overlaps with zero.

ggplot(f2, aes(x, mean, group = treatment, col = treatment)) +
  geom_segment(aes(x = x, y = lower, xend = x, yend = upper), size = .75) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 2.5, col = "white") +
  geom_point(aes(col = treatment), size = 1.5) +
  labs(
    y = "Skin Conductance Reactivity (standardized units)",
    title = "The anxiety treatment significantly increased \n skin conductance reactivity."
  ) +
  scale_color_discrete(name = "Treatment") +
  theme(
    text = element_text(family = "LM Roman 10"),
    panel.background = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.key = element_blank()
  ) +
  geom_text(aes(label = paste("n = ", count)), y = -.3, x = c(1.15, 2, 2.95), family = "LM Roman 10", show.legend = FALSE)
```
Really nice looking plot. Labels/colors are great. As I said above, you wanted to use SCDBradVidManipAll_mean here to get the
numbers that they report in the paper. Not a huge deal though.

**Figure 2.** Means of skin-conductance reactivity by video condition. 95% confidence intervals shown.

Note that the means (and intervals) are quite different than those shown in the original paper. The overall conclusion, however, that the anxiety treatment significantly increased skin conductance reactivity, is no different.

```{r figure3, echo=FALSE, message=FALSE, error=FALSE, cache=TRUE}
### Figure 3.

## Outcome model
# Regress the immigration score on the anxiety dummy, controling for skin (SCL) score and story group.

mod_outcome <- lm(immigration ~ anxiety + skin + story, no_relax)

## Mediator Model
# Regress the skin (SCL) score on the anxiety dummy while controling for story group.
# Note that this includes the "story" variable, which is not included in the main results.

mod_mediator <- lm(skin ~ anxiety + story, no_relax)

## Mediation Analysis
# This calculates the direct, indirect, and total effects with
# 500 simulations (as in the original paper) and boostraps
# the results for model validation.

great explanations of what's going on here. Really useful for you and the reader.

x <- mediate(mod_mediator,
  mod_outcome,
  sims = 500,
  treat = "anxiety",
  mediator = "skin",
  dropobs = TRUE,
  boot = TRUE,
  conf.level = .90
)

# Save the results to a plottable dataframe that includes
# the average estimate, lower and upper 90% confidence bounds,
# and p values.

f3 <- tibble(
  effect = c("ACME", "Direct Effect", "Total Effect"),
  estimate = c(x$d.avg, x$z.avg, x$tau.coef),
  lower = c(x$d.avg.ci[[1]], x$z.avg.ci[[1]], x$tau.ci[[1]]),
  upper = c(x$d.avg.ci[[2]], x$z.avg.ci[[2]], x$tau.ci[[2]]),
  pval = c(x$d.avg.p, x$z.avg.p, x$tau.p),
  y = c(1, 2, 3)
)

# Plot the result. This plot shows the causal mediation effect of physiological
# reactivity on the dependent variable (immigration beliefs).

ggplot(f3, aes(estimate, y, group = effect, col = effect)) +
  geom_segment(aes(x = lower, y = y, xend = upper, yend = y), size = .75) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_point(aes(estimate), size = 2.5, col = "white") +
  geom_point(aes(estimate, col = effect), size = 1.5) +
  labs(
    x = "Causal mediation effect on immigration beliefs",
    title = "The indirect effect of physiological reactivity (ACME) \n is positive and significant."
  ) +
  scale_color_discrete(name = "Effect") +
  annotate("text", x = -.4, y = 1, label = "More negative beliefs >>>",
           col = "black", alpha = .5, size = 4, family = "LM Roman 10") +
  theme(
    text = element_text(family = "LM Roman 10"),
    panel.background = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.key = element_blank()
  )
```

**Figure 3.** *Treatment* is the anxiety video manipulation (compared to the neutral condition), *Mediator* is skin conductance reactivity when answering immigration questions, and *Outcome* is composite variable of immigration preferences. Horizontal lines represent 90% confidence intervals around the mean estimates.

Note that ACME stands for *average causal mediation effect* and that higher coefficients correspond to less favorable views on immigration.

Good explanation and excellent plot. 

\pagebreak

### Extension

In this extension I attempt to confirm the main results of the paper using Bayesian linear models. I generally follow the guidelines provided in @gabry and add additional visualizations from the `bayesplot` package. The models here do reach the same general conclusions as those from the paper: the *Anxiety* treatment parameter is positive and significant in the first-stage model within a 90% confidence interval, and the skin conductance reactivity parameter is positive and significant in the second-stage model within a 95% confidence interval. This lends credence to the authors' argument that physiological reactivity mediates the relationship between induced anxiety and attitudes toward immigrants.

```{r extension_data, echo=FALSE, message=FALSE, error=FALSE, cache=TRUE, results='hide'}
### Extension

# Reproduce the main results but with a Bayesian model, using
# the rstanarm package.

# Regress SC reactivity on Anxiety (excluding the relax group).

mod1 <- stan_glm(skin ~ anxiety, no_relax, family = "gaussian")

# Regress immigration preferences on anxiety, story, and SC reactivity.

mod2 <- stan_glm(immigration ~ skin + anxiety + story, no_relax, family = "gaussian")

# Separate the posterior distributions for the parameters
# so they can be plotted separately.
# Clean the names for ease of use.

draws1 <- as_tibble(mod1) %>%
  clean_names()

draws2 <- as_tibble(mod2) %>%
  clean_names()
```

```{r xplot1, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, cache=TRUE}
# This extension plot shows the regression line generated from the mean (most likely)
# parameter and coefficient values for anxiety treatment regressed on skin conductance.
# It also graphs all the possible slope/intercept combinations from
# the generated posterior distribution of the anxiety parameter.

The comment above is probably something you want to include written in your report. That way it's clear to the reader what is going on
and there is no guessing about what your figure is showing. Something to keep in mind for final project/paper.

ggplot(no_relax, aes(anxiety, skin, col = anxiety)) +
  scale_color_manual(values = c("black", "firebrick1"), name = "Treatment", labels = c("Neutral", "Anxiety")) +
  geom_point(size = 1, position = position_jitter(height = 0.05, width = 0.1)) +
  geom_abline(
    data = draws1, aes(intercept = intercept, slope = anxiety1),
    color = "skyblue", size = 0.1, alpha = 0.05
  ) +
  geom_abline(
    intercept = coef(mod1)[1], slope = coef(mod1)[2],
    color = "deepskyblue4", size = 1
  ) +
  labs(
    x = "Video treatment",
    y = "Skin Conductance Reactivity",
    title = "Anxiety treatment is associated with increased \n skin conductance reactivity."
  ) +
  theme(
    text = element_text(family = "LM Roman 10"),
    panel.background = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
```

```{r xplot2, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, fig.align='center', fig.width=5, fig.height=2.5}
# This plot shows the mean value and 90% confidence interval of the
# anxiety treatment in the first model generated from the
# posterior distribution. I choose 90% here to match with the results
# shown in the original paper.

Same as above. Definitely want this in the report so it's clear what's going on. 

mcmc_areas(draws1, prob = 0.90, pars = "anxiety1", point_est = "mean") +
  labs(
    x = "Anxiety parameter value",
    title = "Posterior distribution estimate of anxiety parameter",
    caption = "Mean and 90% interval shown in blue."
  ) +
  theme(
    text = element_text(family = "LM Roman 10"),
    panel.background = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  geom_segment(aes(x = mean(draws1$anxiety1), y = 0, xend = mean(draws1$anxiety1), yend = 1), size = .15, linetype = "dashed") +
  annotate("text", label = round(mean(draws1$anxiety1), 3), x = mean(draws1$anxiety1) - .1, y = .2, family = "LM Roman 10")
```

Note that the *Anxiety* parameter value is positive and significant within a 90% confidence interval.

```{r xplot3, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, cache=TRUE}
# This extension plot shows the regression line generated from the mean (most likely)
# parameter and coefficient values for immigration opinions regressed on skin conductance
# while holding video and story treatment constant.
# It also graphs all the possible slope/intercept combinations from
# the generated posterior distribution of the skin conductance parameter.

ggplot(no_relax, aes(skin, immigration)) +
  geom_point(size = 1, position = position_jitter(height = 0.05, width = 0.1)) +
  labs(
    x = "Skin Conductance Reactivity",
    y = "Immigration views",
    title = "Skin reactivity mediates views on immigration."
  ) +
  geom_abline(
    data = draws2, aes(intercept = intercept, slope = skin),
    color = "skyblue", size = 0.1, alpha = 0.05
  ) +
  geom_abline(
    intercept = coef(mod2)[1], slope = coef(mod2)[2],
    color = "skyblue4", size = 1
  ) +
  theme(
    text = element_text(family = "LM Roman 10"),
    panel.background = element_blank()
  )
```

```{r xplot4, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, fig.align='center', fig.width=5, fig.height=2.5}
# This plot shows the mean value and 95% confidence interval of the
# skin conductance parameter in the first model generated from the
# posterior distribution. I choose 95% here to match with the results
# shown in the original paper.

mcmc_areas(draws2, prob = 0.95, pars = "skin", point_est = "mean") +
  labs(
    x = "Skin conductance reactivity parameter value",
    title = "Posterior distribution of skin conductance parameter",
    caption = "Mean and 95% interval shown in blue."
  ) +
  theme(
    text = element_text(family = "LM Roman 10"),
    panel.background = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  geom_segment(aes(x = mean(draws2$skin), y = 0, xend = mean(draws2$skin), yend = 1), size = .15, linetype = "dashed") +
  annotate("text", label = round(mean(draws2$skin), 3), x = mean(draws2$skin) + .05, y = .3, family = "LM Roman 10")
```

Note that the skin conductance reactivity parameter value is positive and significant within a 95% confidence interval. This Bayesian model had the same controls as that of the original, namely anxiety manipulation and story condition.

Nice extension overall. The plots provide some nice visuals that add to the claims made in the paper. As said above, you want to say more about your plots in your report so someone who is reading it without seeing the comments knows exactly what's going on
and what each figure means. Nice job!

\pagebreak

### References
