---
title: "Midterm: Physiological Arousal and Political Beliefs"
output:
  pdf_document:
    md_extension: +raw_tex
bibliography: citations.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stargazer)
library(stringr)
library(bookdown)
library(mediation)
library(memisc)
library(extrafont)

# Replication of "Physiological Arousal and Political Beliefs"
# Published in Political Psychology (2014)
# Original authors: Jonathan Renshon, Julia Lee & Dustin Tingley
```

\begin{flushright}
Gabe Walker

March 16, 2019
\end{flushright}

```{r data, echo=FALSE, message=FALSE, error=FALSE}

# Read in the replication file provided on the Harvard Dataverse.
# https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/24318

df <- read_csv("replication_files/anxiety.csv") %>% 
  
# Remove the two observations who have a lot of missing data.
  
  filter(! {is.na(income) & is.na(education)}) %>% 

# Rename and restructure variables for ease of understanding.
# "video" refers to which video the subject was shown.

  mutate(video = factor(case_when(
    anxcond3 == 0 ~ "relax",
    anxcond3 == 1 ~ "neutral",
    anxcond3 == 2 ~ "anxiety"
  ))) %>% 
  
  dplyr::select(-anxcond3) %>% 

# "anxiety" is a dummy for whether the subject was shown the Cliffhanger clip.
    
  mutate(anxiety = factor(anxcond)) %>% 
  dplyr::select(-anxcond) %>% 

# "skin" is the standardized skin-conductance level within each participant.
# See the original paper (p. 7) for a detailed description of standardization.

  mutate(skin = SCDBradSelfReport1_mean) %>% 
  dplyr::select(-SCDBradSelfReport1_mean) %>% 

# "story" is a random dummy for which immigration article the participants read.
  
  mutate(story = factor(storycond)) %>% 
  dplyr::select(-storycond)

```

### Abstract

I partially replicate "Physiological Arousal and Political Beliefs" by @renshon from the [Harvard Dataverse](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/24318). The main regression results (Table 1) appear exactly as in the original paper. I have improved the aesthetics of Figures 2 and 3 with a more minimal and unified style. Note that the results of Figure 2 are different than in the original. Though I find significantly different skin-conductance reactivities for each video  treatment group, the overall conclusions are no different. Figure 3 results are the same as in the paper. For the original replication data, see @data.

\vspace{2em}

**Table 1.** Main Results

\begin{center}

```{r echo=FALSE, message=FALSE, error=FALSE, results='asis'}
### Table 1.

# Subset the data to only include individuals in the
# "neutral" or "anxiety" video treatment groups.

no_relax <- df %>% 
  filter(video != "relax")

# Regress SC reactivity on Anxiety (excluding relax)

mod1 <- lm(skin ~ anxiety, no_relax)

# Regress immigration preferences on anxiety, story, sc reactivity

mod2 <- lm(immigration ~ anxiety + story + skin, no_relax)

# Table should show N, R^2

stargazer(mod1, mod2,
  style = "apsr",
  omit.stat = c("adj.rsq", "ll", "F", "ser"),
  covariate.labels = c("Anxiety Manipulation", "Story Condition", "SC Reactivity while answering questions"),
  dep.var.labels = c("SC Reactivity", "Immigration Preferences"),
  column.sep.width = "5pt",
  float = FALSE,
  header = FALSE)
```

\end{center}

*Note.* Model (1) shows the effect of the treatment (anxiety) on physiological reactivity while Model (2) shows the effects
of physiological reactivity on immigration preferences, controlling for the story condition. Both models includes only
*Neutral* & *Anxiety* conditions (*Relax* condition is excluded). 

Standard errors in parentheses.


```{r echo=FALSE, message=FALSE, error=FALSE, cache=TRUE}
### Figure 2.

# Make three data subsets for each treatment group.
# Remove the NA values so that certain functions (e.g., mean()) work.

relax <- df %>%
  filter(video == "relax") %>%
  na.omit()

neutral <- df %>%
  filter(video == "neutral") %>%
  na.omit()

anxiety <- df %>%
  filter(video == "anxiety") %>%
  na.omit()

# Make a plottable dataframe with the means,
# confidence intervals, and counts for the "skin"
# variable for each group

# Find the means of each group.

means <- c(mean(relax$skin), mean(neutral$skin), mean(anxiety$skin))

# Find the confidence interval sizes for each group.

ci_relax <- 1.96 * sd(relax$skin) / sqrt(length(relax$skin))
ci_neutral <- 1.96 * sd(neutral$skin) / sqrt(length(neutral$skin))
ci_anxiety <- 1.96 * sd(anxiety$skin) / sqrt(length(anxiety$skin))

# Find the upper and lower bounds for each group.

lower <- c(means[1] - ci_relax, means[2] - ci_neutral, means[3] - ci_anxiety)
upper <- c(means[1] + ci_relax, means[2] + ci_neutral, means[3] + ci_anxiety)

# Put the results in a graphable dataframe.

f2 <- tibble(
  treatment = c("Relax", "Neutral", "Anxiety"),
  mean = means,
  lower = lower,
  upper = upper,
  count = c(nrow(relax), nrow(neutral), nrow(anxiety)),
  x = c(1,2,3)
)

# Plot the result.

ggplot(f2, aes(x, mean, group = treatment, col = treatment)) +
  geom_segment(aes(x = x, y = lower, xend = x, yend = upper), size = .75) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 2.5, col = "white") +
  geom_point(aes(col = treatment), size = 1.5) +
  labs(y = "Skin Conductance Reactivity",
       title = "The anxiety treatment significantly increased skin conductance reactivity") +
  scale_color_discrete(name = "Treatment") +
  theme(text = element_text(family = "LM Roman 10"),
        panel.background = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.key = element_blank()) +
  geom_text(aes(label = paste("n = ", count)), y = -.3, x = c(1.15, 2, 2.95), family = "LM Roman 10", show.legend = FALSE)
```

**Figure 2.** Means of skin-conductance reactivity by video condition. 95% confidence intervals shown. (p. 8)

Note that the means (and lines) are quite different than those shown in the original paper. The overall conclusion, however, that the anxiety treatment significantly increased skin conductance reactivity, is no different.

```{r echo=FALSE, message=FALSE, error=FALSE, cache=TRUE}
### Figure 3.

# Subset the data to only include individuals in the
# "neutral" or "anxiety" video treatment groups.

no_relax <- df %>% 
  filter(video != "relax")

## Outcome model
# Regress the immigration score on the anxiety dummy, controling for skin (SCL) score and story group.

mod_outcome <- lm(immigration ~ anxiety + skin + story, no_relax)

## Mediator Model
# Regress the skin (SCL) score on the anxiety dummy while controling for story group.
# Note that this includes the "story" variable, which is not in the main results.

mod_mediator <- lm(skin ~ anxiety + story, no_relax)

## Mediation Analysis

x <- mediate(mod_mediator, 
                     mod_outcome, 
                     sims = 500, 
                     treat = "anxiety", 
                     mediator = "skin", 
                     dropobs = TRUE, 
                     boot = TRUE, 
                     conf.level = .90)

# Save the results to a plottable dataframe.

f3 <- tibble(
  effect = c("ACME", "Direct Effect", "Total Effect"),
  estimate = c(x$d.avg, x$z.avg, x$tau.coef),
  lower = c(x$d.avg.ci[[1]], x$z.avg.ci[[1]], x$tau.ci[[1]]),
  upper = c(x$d.avg.ci[[2]], x$z.avg.ci[[2]], x$tau.ci[[2]]),
  pval = c(x$d.avg.p, x$z.avg.p, x$tau.p),
  y = c(1,2,3)
)

# Plot the result.

ggplot(f3, aes(estimate, y, group = effect, col = effect)) +
  geom_segment(aes(x = lower, y = y, xend = upper, yend = y), size = .75) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_point(aes(estimate), size = 2.5, col = "white") +
  geom_point(aes(estimate, col = effect), size = 1.5) +
  labs(x = "Causal mediation effect",
       title = "The indirect effect of physiological reactivity (ACME) is positive and significant") +
  scale_color_discrete(name = "Effect") +
  theme(text=element_text(family = "LM Roman 10"),
        panel.background = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.key = element_blank())
```

**Figure 3.** *Treatment* is anxiety video manipulation (compared to neutral condition), *Mediator* is skin conductance reactivity when answering immigration questions, *Outcome* is composite variable of immigration preferences. Horizontal lines represent 90% confidence intervals for estimates. (p. 11)

N.B., ACME stands for *average causal mediation effect*.

### Extension


### References
